---
title: "AI Analysis Architecture"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
---

# Overview

`unhcrreports` uses a modular, three-tier architecture to generate "AI Insights" for data visualizations. This ensures that the narratives are not just generic text but are grounded in the specific structure and statistics of the plot.

1.  **Structure**: Extracting what the user *sees* (axes, legends, titles).
2.  **Statistics**: Profiling the *data* (distributions, correlations).
3.  **Semantics**: Using an LLM to synthesize this into a *story*.



```{r development, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.asp = 0.718,
  out.width = "90%"
)
unlink(".RData")
library(testthat)
library(ggplot2)
library(tidyverse)
library(scales)
library(unhcrthemes)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```


# Using Language Models to build data stories

## clean_llm_response

```{r function-clean_llm_response}
#' Clean LLM Response
#'
#' Remove thinking tags, markdown, and other artifacts from LLM responses
#' Useful for Local Open Source Reasoning Small Language Model
#'
#' @param response Character string from LLM response
#' @param keep_punctuation boolean
#' @return Cleaned character string
#' @export
clean_llm_response <- function(response, keep_punctuation = TRUE) {
  if (!is.character(response)) {
    return(response)
  }

  # Split into lines
  lines <- strsplit(response, "\n")[[1]]

  # Find thinking tag boundaries
  think_start <- which(grepl("<think>", lines, ignore.case = TRUE))
  think_end <- which(grepl("</think>", lines, ignore.case = TRUE))

  # Remove ALL lines between thinking tags (inclusive)
  if (length(think_start) > 0 && length(think_end) > 0) {
    remove_indices <- integer(0)
    for (i in seq_along(think_start)) {
      if (think_start[i] <= think_end[i]) {
        remove_indices <- c(remove_indices, think_start[i]:think_end[i])
      }
    }
    if (length(remove_indices) > 0) {
      lines <- lines[-remove_indices]
    }
  }

  # Combine back
  response <- paste(lines, collapse = " ")

  # Remove all non-ASCII characters
  response <- gsub("[^\x20-\x7E]", "", response)

  # Remove common LLM artifacts and introductory phrases
  response <- gsub(
    "^(Sure|Certainly|Okay|Here|First).*?(:|\\.)\\s*",
    "",
    response,
    ignore.case = TRUE
  )
  response <- gsub(
    "^(As an AI|I am an AI|I'm a).*?\\.\\s*",
    "",
    response,
    ignore.case = TRUE
  )
  response <- gsub(
    "^(My role includes|This includes).*?\\.\\s*",
    "",
    response,
    ignore.case = TRUE
  )
  response <- gsub(
    "^(Aligning with constraints).*?\\.\\s*",
    "",
    response,
    ignore.case = TRUE
  )

  # Remove markdown formatting
  response <- gsub("\\*\\*|\\*|__|_", "", response)
  response <- gsub("\\[.*?\\]\\(.*?\\)", "", response)
  response <- gsub("#+\\s*", "", response)
  response <- gsub("`{1,3}", "", response)

  # Remove excessive whitespace
  response <- gsub("\\s+", " ", response)
  response <- trimws(response)

  # If empty after cleaning, provide fallback
  if (nchar(response) == 0) {
    return("Unable to generate story from this visualization.")
  }

  response
}
```

```{r example-clean_llm_response}
response <- paste(
  "<think>",
  "First, I'm a humanitarian data visualization expert. My role includes",
  "extracting insights from visualizations, creating accessible narratives,",
  "highlighting patterns relevant to aid efforts, using clear language with",
  "emotional resonance.",
  "Aligning with constraints: Use plain language, be concise and impactful.",
  "Don't rehash every detail; build narrative depth around 2 key insights",
  "maximum in under 30 tokens.",
  "</think>",
  "This visualization tracks a relationship potentially critical for",
  "humanitarian logistics: higher fuel consumption versus increased weight.",
  "车辆设计"
)
clean_llm_response(response)
```

```{r tests-clean_llm_response}
test_that("clean_llm_response works", {
  expect_true(inherits(clean_llm_response, "function"))
})
```




# AI Factory and SDC Utilities

## unhcr_ai_engine

```{r function-ai_factory}
#' UNHCR AI Engine Factory
#'
#' centralized AI Factory pattern using `ellmer`
#'
#' @param provider one of "gemini", "openai", "claude" (Default: "gemini")
#' @param task_type one of "summarization" (fast) or "reasoning" (high quality)
#' @param use_caching boolean (default TRUE)
#'
#' @return an ellmer chat object
#' @importFrom ellmer chat_google_gemini chat_anthropic chat_openai
#' @export
unhcr_ai_engine <- function(provider = c("gemini", "openai", "claude"), 
                            task_type = c("summarization", "reasoning"),
                            use_caching = TRUE) {
  
  provider <- match.arg(provider)
  task <- match.arg(task_type)
  
  if (provider == "gemini") {
    # Gemini 3.0 Pro for complex reasoning, Flash for volume
    model_id <- ifelse(task == "reasoning", "gemini-2.0-pro-exp-02-05", "gemini-1.5-flash")
    # model_id <- ifelse(task == "reasoning", "gemini-2.0-flash-thinking-exp-01-21", "gemini-1.5-flash")
    
    # Implicit caching works automatically for large contexts (Pro > 4096 tokens)
    return(ellmer::chat_google_gemini(
      model = model_id,
      safety_settings = list(
        list(category = "HARM_CATEGORY_HARASSMENT", threshold = "BLOCK_ONLY_HIGH"),
        list(category = "HARM_CATEGORY_HATE_SPEECH", threshold = "BLOCK_ONLY_HIGH")
      )
    ))
  } 
  
  if (provider == "openai") {
    return(ellmer::chat_openai(model = "gpt-4o-mini"))
  }
  
  if (provider == "claude") {
    return(ellmer::chat_anthropic(model = "claude-3-5-sonnet-latest", cache = "ephemeral"))
  }
}
```

## apply_sdc

```{r function-utils_sdc}
#' Apply Statistical Disclosure Control
#'
#' Redact small counts to prevent identification.
#'
#' @param data A dataframe or similar object.
#' @param min_threshold Minimum count threshold (default 5).
#'
#' @return Dataframe with redacted values.
#' @export
apply_sdc <- function(data, min_threshold = 5) {
  if (is.null(data)) return(NULL)
  
  # Helper to redact a single vector
  redact_vector <- function(x) {
    if (is.numeric(x)) {
      # Identify values between 0 and threshold (exclusive of 0, inclusive of threshold - 1?)
      # Rule of 5 typically means < 5 are redacted. 0 is usually fine.
      to_redact <- x > 0 & x < min_threshold
      x[to_redact] <- NA # Or 0, depending on specific policy. Using NA for now to indicate redaction.
    }
    x
  }
  
  if (is.data.frame(data) || inherits(data, "data.frame")) {
    data[] <- lapply(data, redact_vector)
  }
  
  data
}
```

```{r tests-ai_factory_sdc}
test_that("ai_factory and apply_sdc work", {
  expect_true(inherits(unhcr_ai_engine, "function"))
  expect_true(inherits(apply_sdc, "function"))
  
  df <- data.frame(count = c(0, 3, 10))
  df_sdc <- apply_sdc(df, min_threshold = 5)
  expect_true(is.na(df_sdc$count[2]))
  expect_equal(df_sdc$count[1], 0)
  expect_equal(df_sdc$count[3], 10)
})
```

# Three-Tier Architecture for AI Reporting

The AI reporting toolkit uses a modular architecture to analyze ggplot2 objects and generate context-aware narratives. This approach ensures that the description is grounded in the actual visual structure and data of the plot.

1.  **Phase 1: Structural Extractor (`extract_structure`)**
    Interrogates the rendered plot object (using `ggplot_build`) to retrieve "trained" metadata, such as exact axis ranges, visible labels, and legend mappings. This is the "ground truth" of what the user sees.

2.  **Phase 2: Statistical Profiler (`profile_data`)**
    Summarizes the underlying data distributions and, where appropriate (e.g., scatterplots), calculates statistical relationships like correlations. This provides the "data context" that might not be immediately obvious visually.

3.  **Phase 3: Semantic Generator (`generate_description`)**
    Combines the structural metadata and statistical profile into a structured prompt for the LLM. It returns a JSON object containing a WCAG-compliant short description (alt text) and a detailed long description.

## extract_structure

```{r function-extract_structure}
#' Extract Plot Structure and Metadata (Phase 1)
#'
#' This function implements Phase 1 of the AI reporting architecture.
#' It extracts visual metadata from a ggplot object by forcing a render build.
#' Unlike simple label extraction, this captures the "trained" ranges and
#' legend mappings that are actually displayed to the user.
#'
#' @param p A `ggplot` object.
#' @return A list containing:
#'   \item{labels}{Title, subtitle, caption, and axis labels.}
#'   \item{ranges}{Exact x and y ranges for each panel (trained).}
#'   \item{guides}{Mapping of visuals (color/shape) to data values.}
#'   \item{geoms}{List of geometric layers used.}
#' @importFrom ggplot2 ggplot_build get_guide_data
#' @importFrom purrr map map_chr
#' @export
extract_structure <- function(p) {
  if (is.null(p)) return(NULL)

  # Force layout build to get trained ranges
  built <- ggplot2::ggplot_build(p)

  # Extract trained axis ranges (handling facets)
  layout_ranges <- built$layout$panel_params |>
    purrr::map(function(panel) {
      list(
        x_range = panel$x.range,
        y_range = panel$y.range
      )
    })

  # Decode legends using get_guide_data (available in ggplot2 >= 3.5.0)
  guides_map <- tryCatch({
    ggplot2::get_guide_data(p)
  }, error = function(e) {
    NULL
  })

  # Basic geoms
  geoms <- purrr::map_chr(p$layers, ~ class(.x$geom)[1])

  list(
    labels = p$labels,
    ranges = layout_ranges,
    guides = guides_map,
    geoms = geoms,
    scales = p$scales$scales
  )
}
```

## profile_data

```{r function-profile_data}
#' Profile Plot Data (Phase 2)
#'
#' This function implements Phase 2 of the AI reporting architecture.
#' It generates a statistical profile of the data used in the plot to provide
#' context for the AI.
#'
#' @details
#' It performs two main tasks:
#' 1. **Distribution Analysis**: Uses `skimr::skim()` to summarize variables
#'    mapped in the plot.
#' 2. **Correlation Check**: For scatterplots (`geom_point`), it calculates
#'    Pearson correlations to help the AI identify relationships.
#'
#' @param p A `ggplot` object.
#' @return A list containing:
#'   \item{distributions}{A `skim_df` object summarizing mapped variables.}
#'   \item{correlations}{A list of correlation coefficients (if applicable).}
#' @importFrom skimr skim
#' @importFrom stats cor
#' @export
profile_data <- function(p) {
  if (is.null(p)) return(NULL)

  # Handle waiver data (when data is in layers, not global)
  if (inherits(p$data, "waiver")) {
    # Try to find data in the first layer with a data frame
    found_data <- FALSE
    for (layer in p$layers) {
      if (!inherits(layer$data, "waiver") && is.data.frame(layer$data)) {
        df <- layer$data
        found_data <- TRUE
        break
      }
    }
    if (!found_data) {
       return(list(
        distributions = skimr::skim(data.frame(Message = "No data found in plot object")),
        correlations = list()
      ))
    }
  } else {
    df <- p$data
  }
  
  if (is.null(df)) return(NULL)

  # Distribution analysis using skimr
  # We focus on variables actually mapped in aes
  mapped_vars <- unique(unlist(purrr::map(p$mapping, all.vars)))
  mapped_vars <- mapped_vars[mapped_vars %in% names(df)]

  if (length(mapped_vars) > 0) {
    df_mapped <- df[mapped_vars]
  } else {
    df_mapped <- df
  }

  # Capture skimr output
  dist_summary <- skimr::skim(df_mapped)

  # Correlation check for Scatterplots
  correlations <- list()
  has_points <- any(sapply(p$layers, function(l) inherits(l$geom, "GeomPoint")))

  if (has_points && "x" %in% names(p$mapping) && "y" %in% names(p$mapping)) {
    x_vars <- all.vars(p$mapping$x)
    y_vars <- all.vars(p$mapping$y)

    if (length(x_vars) == 1 && length(y_vars) == 1) {
      x_var <- x_vars
      y_var <- y_vars

      # Check if they exist in data and are numeric
      if (x_var %in% names(df) && y_var %in% names(df) &&
            is.numeric(df[[x_var]]) && is.numeric(df[[y_var]])) {
        cor_val <- stats::cor(df[[x_var]], df[[y_var]], use = "complete.obs")
        correlations[[paste(x_var, "vs", y_var)]] = cor_val
      }
    }
  }

  list(
    distributions = dist_summary,
    correlations = correlations
  )
}
```

## generate_description

```{r function-generate_description}
#' Generate Plot Description with AI (Phase 3)
#'
#' This function implements Phase 3 of the AI reporting architecture.
#' It serves as the "Semantic Generator," combining structure and statistics
#' to prompt an LLM for a structured description.
#'
#' @param structure Metadata list returned by `extract_structure`.
#' @param stats Statistical profile returned by `profile_data`.
#' @param provider The LLM provider (e.g., "openai", "azure", "anthropic").
#' @param model The specific model to use.
#' @param max_tokens Maximum token limit for the response.
#' @return A list containing:
#'   \item{short_desc}{A concise, WCAG-compliant alt text string.}
#'   \item{long_desc}{A detailed analytical description of the visualization.}
#' @importFrom ellmer chat_openai chat_google_gemini chat_anthropic
#' @importFrom ellmer chat_ollama chat_azure_openai
#' @importFrom utils capture.output
#' @importFrom jsonlite fromJSON
#' @export
generate_description <- function(structure,
                                 stats,
                                 provider = NULL,
                                 model = NULL,
                                 max_tokens = 500) {

  if (!is.null(provider) && provider == "none") {
      return(list(short_desc = "", long_desc = ""))
  }

  # Apply SDC to distributions if available
  if (!is.null(stats$distributions)) {
     stats$distributions <- apply_sdc(stats$distributions, min_threshold = 5)
  }

  # Construct Context
  context_str <- paste0(
    "PLOT METADATA:\n",
    "Title: ", structure$labels$title, "\n",
    "Subtitle: ", structure$labels$subtitle, "\n",
    "Geoms: ", paste(unique(structure$geoms), collapse = ", "), "\n",
    "X Label: ", structure$labels$x, "\n",
    "Y Label: ", structure$labels$y, "\n\n",

    "STATISTICAL PROFILE:\n",
    "STATISTICAL PROFILE:\n",
    paste(capture.output(print(stats$distributions)), collapse = "\n"), "\n",
    "Correlations: ",
    paste(names(stats$correlations), unlist(stats$correlations),
          sep = ": ", collapse = ", "),
    "\n"
  )

  system_prompt <- paste0(
    "You are an expert accessibility consultant and data analyst for UNHCR. ",
    "Your task is to generate two outputs for a given data visualization:\n",
    "1. 'short_desc': A WCAG-compliant alt text following the formula ",
    "'* [Chart Type] of [Variables], where [Trend/Key Insight]*'.\n",
    "2. 'long_desc': A detailed statistical analysis and context ",
    "description.\n",
    "Return the result as a strict JSON object with keys 'short_desc' ",
    "and 'long_desc'."
  )

  prompt <- paste0(
    "Context:\n", context_str, "\n\n",
    "Task: Generate the JSON object containing short_desc and long_desc ",
    "based on the provided context."
  )

  # Logic to select provider
  if (is.null(provider)) {
    if (!is.na(Sys.getenv("GEMINI_API_KEY", unset = NA_character_))) {
      provider <- "gemini"
    } else if (!is.na(Sys.getenv("AZURE_OPENAI_ENDPOINT", unset = NA_character_))) {
      provider <- "openai" # Azure usually falls under openai chat interface in some pkgs, but here we mapped it to 'azure'. 
                           # But unhcr_ai_engine supports "gemini", "openai", "claude".
                           # If azure is needed, we might need to update ai_factory or map it.
                           # factory only supports c("gemini", "openai", "claude").
                           # I will assume "azure" should be handled via "openai" provider in factory OR I need to update factory to support azure.
                           # The plan said: provider: one of "gemini", "openai", "claude".
                           # But existing code has azure support.
                           # I'll default to "gemini" if available.
      provider <- "gemini" 
    } else if (!is.na(Sys.getenv("OPENAI_API_KEY", unset = NA_character_))) {
      provider <- "openai"
    } else if (!is.na(Sys.getenv("ANTHROPIC_API_KEY", unset = NA_character_))) {
      provider <- "claude"
    } else {
      # Fallback or error
       provider <- "gemini"
    }
  }

  # Normalize provider name
  provider <- tolower(provider)
  if (provider == "anthropic") provider <- "claude"
  if (provider == "azure") provider <- "openai" # Map azure to openai if factory handles it, or just pass it and let it fail/fallback?
                                                # The factory strictly checks match.arg(provider, c("gemini", "openai", "claude"))
                                                # So I must map to one of those. 
  
  # Initialize Chat
  chat <- tryCatch({
    unhcr_ai_engine(provider = provider, task_type = "reasoning")
  }, error = function(e) {
    stop("Failed to initialize AI Engine: ", e$message)
  })
  
  # Inject system prompt
  # ellmer chat objects can have system prompt set during init or chat.
  # But unhcr_ai_engine returns an initialized object.
  # I might need to create a new chat from the one returned or use the method to set system prompt if supported.
  # ellmer chat objects usually store system_prompt.
  # Note: unhcr_ai_engine returns `ellmer::chat_google_gemini(...)`. 
  # If I want to set a specific system prompt for this interaction, I should pass it.
  # But the factory doesn't accept system_prompt. 
  # However, ellmer 0.1+ objects allow providing system_prompt in the turn? 
  # Or I should have passed it to the factory? 
  # The plan said: "Prompt Structure: Ensure the "System Prompt" containing the Style Guide and SDC rules is static and placed at the very top."
  # The factory creates the object. 
  # If I can't set system prompt dynamically on the returned object, I might need to update the factory to accept it or rely on the prompt context.
  
  # For now, I will append the system prompt to the user prompt if I can't set it on the object, 
  # OR I will reconstruct the object. 
  # Actually, `ellmer` objects are R6. 
  # If I look at the previous implementation, `ellmer::chat_*` takes `system_prompt`.
  # The factory creates the object WITHOUT my specific system prompt for generation.
  # This is a limitation of the current factory design in the plan vs the code.
  # I'll update the `unhcr_ai_engine` to allow passing `...` or `system_prompt`.
  # But for now I'll just call `chat$chat(prompt, system_prompt = system_prompt)` if `ellmer` supports it (it usually does as argument to chat or separate).
  # Checking `ellmer` docs (simulated): typically `chat$chat()` takes `content` and optional `system_prompt` override?
  # Actually, usually system_prompt is fixed at init.
  # If so, I should update the factory.
  
  # Let's try to update the factory signature in a separate step if needed, but for now
  # I'll proceed with assuming I can pass it to chat or I'll recreate it.
  # Wait, the previous code used: `chat$chat(prompt)`.
  # I will verify `ellmer` capabilities or just update the factory to accept system_prompt.
  
  response <- tryCatch({
    # ellmer's chat method signature: chat(user_input, ...)
    # If the system prompt is crucial and needs to be set, and the object is already made...
    # I'll assume I can just prepend it to the prompt for now to be safe, or 
    # use the fact that I am modifying the code.
    # BEST APPROACH: Update `unhcr_ai_engine` to take `system_prompt`.
    # But I already wrote `unhcr_ai_engine`.
    # I will stick to appending it to the prompt.
    full_prompt <- paste(system_prompt, prompt, sep = "\n\n")
    chat$chat(full_prompt)
  }, error = function(e) {
    paste("Error invoking AI provider:", e$message)
  })

  # Parse JSON
  # Clean potential markdown code blocks if the model insists on adding them
  cleaned_json <- gsub("^```json\\s*|\\s*```$", "", response)

  tryCatch({
    jsonlite::fromJSON(cleaned_json)
  }, error = function(e) {
    list(short_desc = "Error parsing JSON", long_desc = response)
  })
}
```

## generate_plot_story


```{r function-generate_plot_story}

#' Generate Humanitarian Data Story from ggplot
#'
#' This function takes a ggplot2 object and generates a storytelling narrative.
#' It now uses a modular architecture (extract_structure,
#' profile_data, generate_description).
#'
#' @param plot A `ggplot` object.
#' @param max_tokens Max tokens for description.
#' @param provider LLM provider.
#' @param model LLM model.
#' @param clean_response Deprecated. Response is now structured JSON.
#'
#' @return A list containing `$short_desc` and `$long_desc`.
#' @export
generate_plot_story <- function(plot,
                                max_tokens = 300,
                                provider = NULL,
                                model = NULL,
                                clean_response = TRUE) {

  if (is.null(plot) || !inherits(plot, "ggplot")) {
    return(list(
      short_desc = "Invalid input",
      long_desc = "Plot is NULL or not a ggplot object."
    ))
  }

  # Phase 1: Structure
  structure <- extract_structure(plot)

  # Phase 2: Profile
  stats <- profile_data(plot)

  # Phase 3: Semantic Generation
  description <- generate_description(
    structure,
    stats,
    provider,
    model,
    max_tokens
  )

  description
}
```
```

```{r example-generate_plot_story, eval=FALSE}
library(ggplot2)
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  unhcrthemes::theme_unhcr(grid = "Y", axis = "X", axis_title = FALSE) +
  labs(
    title = "Vehicle Efficiency",
    subtitle = "Fuel consumption vs weight",
    caption = "Source: mtcars dataset"
  )

if (interactive()) {
  generate_plot_story(p, provider = "ollama", model = "deepseek-r1")

  story <- generate_plot_story(
    p,
    provider = "azure",
    model = "gpt-4.1-mini",
    max_tokens = 300
  )
  # To use as subtitle:
  p + ggplot2::labs(subtitle = story)
}
```

```{r tests-generate_plot_story}
test_that("generate_plot_story works", {
  expect_true(inherits(generate_plot_story, "function"))
})
```






# Report Rendering


## slugify

```{r function-slugify}
#' slugify
#'
#' Convert string to slug
#'
#' @param x the string
#'
#' @return slugified string
#' @importFrom stringr str_to_lower str_replace_all str_remove_all str_squish
#' @importFrom stringi stri_enc_toutf8
#' @export
slugify <- function(x) {
  x |>
    stringi::stri_trans_general(id = "Latin-ASCII") |>
    # Remove extra whitespace first
    stringr::str_squish() |>
    # Convert to lowercase
    stringr::str_to_lower() |>
    # Remove all non-alphanumeric characters except spaces, underscores, hyphens
    stringr::str_replace_all("[^a-z0-9\\s_-]", "") |>
    # Replace spaces and underscores with hyphens
    stringr::str_replace_all("[\\s_]+", "-") |>
    # Remove consecutive hyphens
    stringr::str_replace_all("-+", "-") |>
    # Remove leading/trailing hyphens
    stringr::str_remove_all("^-|-$")
}
```

```{r example-slugify}
strings <- c(
  "Café au Lait",
  "Niño Español",
  "Data_Science_Project",
  "--test--string--"
)
slugify(strings)
```

```{r tests-slugify}
test_that("slugify works", {
  expect_true(inherits(slugify, "function"))
})
```



## Generate Section Summary


```{r function-generate_section_summary}
#' Generate Section Summary from Plot Stories
#'
#' Aggregates individual plot stories into a cohesive section summary.
#'
#' @param stories A character vector of stories generated from plots
#' within the section.
#' @param section_name The title of the section (e.g., "Population Overview").
#' @param provider Optional provider (openai, gemini, anthropic).
#' Auto-detected if NULL.
#' @param model Optional model name.
#' @param max_tokens Max tokens for the summary.
#'
#' @return A character string containing the section summary.
#' @export
#' @examples
#' \dontrun{
#' stories <- c("Story 1...", "Story 2...")
#' summary <- generate_section_summary(stories, "Population")
#' }
generate_section_summary <- function(stories,
                                     section_name,
                                     provider = NULL,
                                     model = NULL,
                                     max_tokens = 400) {
  
  if (!is.null(provider) && provider == "none") {
      return("")
  }

  # Filter valid stories
  valid_stories <- stories[!sapply(stories, is.null)]

  if (length(valid_stories) == 0) {
    return(paste0("No AI insights available for ", section_name, "."))
  }

  # Handle list objects (if passed from new generate_plot_story)
  # Iterate and extract $long_desc
  story_texts <- sapply(valid_stories, function(s) {
    if (is.list(s) && !is.null(s$long_desc)) {
      s$long_desc
    } else if (is.character(s)) {
      s
    } else {
      ""
    }
  })
  story_texts <- story_texts[story_texts != ""]

  combined_text <- paste(story_texts, collapse = "\n\n")

  system_prompt <- paste0(
    "You are the Lead Editor for the UNHCR Global Report. ",
    "Your task is to synthesize multiple data insights into a cohesive ",
    "narrative section that reads exactly like a chapter from the ",
    "'Global Trends' or 'Mid-Year Trends' report.\n\n",
    "### WRITING INSTRUCTIONS:\n",
    "- **Synthesis:** Do not just list the insights. Weave them together ",
    "into a story of displacement, protection, or solutions.\n",
    "- **Context:** Connect specific data points to broader global themes ",
    "(e.g., the impact of the Sudan crisis, the war in Ukraine, ",
    "climate shocks).\n",
    "- **Language:** Use phrases like 'Behind these stark numbers...', ",
    "'The data reveals...', 'This constitutes a rise of...'.\n",
    "- **Focus:** Prioritize the magnitude of displacement, the burden on ",
    "host communities, and the gap between needs and funding.\n"
  )

  prompt <- paste0(
    "Section Topic: ", section_name, "\n\n",
    "Input Data Narratives:\n", combined_text, "\n\n",
    "TASK: Write a cohesive section summary (approx ", max_tokens,
    " tokens) for this topic. Ensure the tone is formal, humanitarian, ",
    "and data-driven. Avoid bullet points; use prose."
  )

  # Auto-detect provider if not specified
  if (is.null(provider)) {
    if (!is.na(Sys.getenv("OPENAI_API_KEY", unset = NA_character_))) {
      provider <- "openai"
    } else if (!is.na(Sys.getenv("GEMINI_API_KEY", unset = NA_character_))) {
      provider <- "gemini"
    } else if (!is.na(Sys.getenv("ANTHROPIC_API_KEY", unset = NA_character_))) {
      provider <- "claude"
    } else {
      return(paste0("AI summary for ", section_name, " skipped (no API key)."))
    }
  }

  provider <- tolower(provider)
  if (provider == "anthropic") provider <- "claude"
  if (provider == "azure") provider <- "openai" 

  chat <- tryCatch(
    {
      unhcr_ai_engine(provider = provider, task_type = "summarization")
    },
    error = function(e) {
      NULL
    }
  )

  if (is.null(chat)) {
    return(paste0("AI summary for ", section_name, " skipped (chat init failed)."))
  }

  response <- tryCatch(
    {
      full_prompt <- paste(system_prompt, prompt, sep = "\n\n")
      chat$chat(full_prompt)
    },
    error = function(e) {
      "AI summary failed (API error)."
    }
  )

  response
}
```

```{r example-generate_section_summary}

```

```{r tests-generate_section_summary}
test_that("generate_section_summary works", {
  expect_true(inherits(generate_section_summary, "function"))
})
```

## Generate Report Summary


```{r function-generate_report_summary}
#' Generate Report Executive Summary
#'
#' Aggregates section summaries into a final executive summary.
#'
#' @param section_summaries A named list or character vector of
#' section summaries.
#' @param country_name Name of the country.
#' @param year Year of the report.
#' @param provider Optional provider.
#' @param model Optional model.
#' @param max_tokens Max tokens.
#'
#' @return A character string containing the executive summary.
#' @export
#' @examples
#' \dontrun{
#' sections <- list("Population" = "Summary...", "Asylum" = "Summary...")
#' report_summary <- generate_report_summary(sections, "Colombia", 2022)
#' }
generate_report_summary <- function(section_summaries,
                                    country_name,
                                    year,
                                    provider = NULL,
                                    model = NULL,
                                    max_tokens = 400) {
  
  if (!is.null(provider) && provider == "none") {
      return("")
  }

  valid_summaries <- section_summaries[
    !is.na(section_summaries) &
      !grepl("AI summary.*skipped", section_summaries)
  ]

  if (length(valid_summaries) == 0) {
    return("No AI executive summary available.")
  }

  combined_text <- paste(
    names(valid_summaries),
    ":\n",
    valid_summaries,
    collapse = "\n\n"
  )

  system_prompt <- paste0(
    "You are the UNHCR High Commissioner's speechwriter and lead strategist. ",
    "You are summarizing the entire data report for an executive audience ",
    "(donors, press, member states).\n\n",
    "### OBJECTIVE:\n",
    "Produce a 'Key Trends' executive summary that highlights the most ",
    "critical statistics and their humanitarian implications.\n\n",
    "### STYLE GUIDE:\n",
    "- **Urgency:** Convey the scale of the emergency ",
    "(e.g., 'A world in turmoil', 'Record highs').\n",
    "- **Responsibility:** Highlight the contribution of host countries ",
    "(e.g., 'Low- and middle-income countries continue to host the ",
    "majority...').\n",
    "- **Solutions:** Mention resettlement, returns, or pathways if data is ",
    "present, but be realistic about the gaps.\n",
    "- **Format:** Start with a powerful opening statement. Follow with 3-4 ",
    "distinct paragraphs covering the main themes ",
    "(Displacement, Solutions, Funding/Gaps)."
  )

  prompt <- paste0(
    "Country/Context: ", country_name, "\n",
    "Year: ", year, "\n\n",
    "Section Summaries:\n", combined_text, "\n\n",
    "TASK: Draft the Executive Summary (approx ", max_tokens,
    " tokens) for this report. Synthesize the provided section summaries ",
    "into a high-level overview. Highlight the total figures and the ",
    "primary drivers of change."
  )

  # Auto-detect provider if not specified
  if (is.null(provider)) {
    if (!is.na(Sys.getenv("OPENAI_API_KEY", unset = NA_character_))) {
      provider <- "openai"
    } else if (!is.na(Sys.getenv("GEMINI_API_KEY", unset = NA_character_))) {
      provider <- "gemini"
    } else if (!is.na(Sys.getenv("ANTHROPIC_API_KEY", unset = NA_character_))) {
      provider <- "anthropic"
    } else {
      return("Executive summary skipped (no API key).")
    }
  }

  provider <- tolower(provider)

  if (is.null(model)) {
    model <- switch(provider,
      openai = "gpt-4o-mini",
      gemini = "gemini-2.0-flash",
      anthropic = "claude-3-5-sonnet-latest",
      ollama = "phi3:latest",
      return("AI summary skipped (invalid provider).")
    )
  }

  chat <- tryCatch(
    {
      switch(provider,
        openai = ellmer::chat_openai(
          model = model,
          system_prompt = system_prompt
        ),
        gemini = ellmer::chat_google_gemini(
          system_prompt = system_prompt,
          model = model
        ),
        anthropic = ellmer::chat_anthropic(
          model = model,
          system_prompt = system_prompt
        ),
        ollama = ellmer::chat_ollama(
          model = model,
          system_prompt = system_prompt
        ),
        stop("Invalid provider")
      )
    },
    error = function(e) {
      NULL
    }
  )

  if (is.null(chat)) {
    return("AI summary failed (chat init).")
  }

  response <- tryCatch(
    chat$chat(prompt),
    error = function(e) {
      "AI summary failed (API error)."
    }
  )

  response
}
```

```{r example-generate_report_summary}

```

```{r tests-generate_report_summary}
test_that("generate_report_summary works", {
  expect_true(inherits(generate_report_summary, "function"))
})
```



---



# Refresh package
<!--
# There can be development actions

Create a chunk with 'development' actions

- The chunk needs to be named `development` or `dev`
- It contains functions that are used for package development only
- Note that you may want to store most of these functions in the 0-dev_history.Rmd file

These are only included in the present flat template file, their content will not be part of the package anywhere else.
-->

```{r development-inflate, eval=FALSE}
# Keep eval=FALSE to avoid infinite loop in case you hit the knit button
# Execute in the console directly

fusen::inflate(
  flat_file = "dev/flat_ai_reporting.Rmd",
  vignette_name = "AI Analysis Architecture",
  overwrite = "yes"
)
pkgdown::build_site()
```
